---
title: "Data Skills Share Presentation - Assessing Black Homeownership Decline in SMCs"
author: "Housing Solutions Lab (Carl Hedman - cgh294@nyu.edu)"
date: "12/03/2021"
output:
  html_document: default
  pdf_document: default
---


```{r load-packages-set-options,results='hide',message = FALSE, warning=FALSE, include=TRUE}
#### Add libraries and formats
#if (system.file(package = "inlmisc", lib.loc = .libPaths()) == "")
 # utils::install.packages("inlmisc", dependencies = TRUE)
library(tidyverse) # general data manipulation and graphing
library(reshape2)
library(lubridate) # format dates
library(scales) # number formatting
library(janitor) # data cleaning
library(hrbrthemes)
library(gt) # formatting tables
library(tidycensus) # pulling in census data
library(jsonlite)
library(ggplot2) # making graphs
library(acssf)
library(ggrepel) # Editing graph lables (we don't use this yet)
library(leaflet) #Creates maps
library(tigris) #Pull Census shapefiles
library(leafpop)# (confirm we need this)
library(plotly)# interactive ggplot for mouseover
library(leaflet.extras)
library(leaflegend)

options(tigris_class = "sf")


```

<br>

#### Background and Methodology
We draw on the 2000 decennial census and 5-year American Community Survey (ACS) data from 2015-2019 to compare indicators for U.S. Census-designated places in all 50 states and D.C. with a population of 50,000 or more (which we will refer to as 2019 for ease of presentation). While not all Census-designated places with populations of 50,000 or greater are cities in terms of their governmental structure, we define them as “cities” for the purposes of this analysis. Based on cities’ 2019 ACS populations, small and midsize cities are defined as places with 50,000 to 499,999 residents, and large cities as places with 500,00 or more residents. We exclude cities established after the 2000 Census and those that merged with their county jurisdiction (e.g., Louisville, KY) over the period. Our final sample includes 782 small and midsize cities.

We utilize Bureau of Labor Statistics' Consumer Price Index (CPI) factor to normalize rents between 2000 and 2019. 

```{r, echo=TRUE,fig.keep='all', results='hide', warning=FALSE, include=TRUE}
### CPI (Consumer Price Index) to adjust for inflation between 2000 and 2019
#Add inflation factor: 2000 rents in 2019 dollars (source: https://www.bls.gov/data/inflation_calculator.htm)
CPI <- 1.49118 # Change the CPI to represent the studied years using the bls inflation calculator. 

### Census API Key
# Input your census API key below to access the census data. You can sign-up for a key here:https://api.census.gov/data/key_signup.html
census_api_key = "c32dfbf7d25fe9558fd11bf021780457970f96ff"

```

<br>
<br>

```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=TRUE}
# Searching census codes for 2015-19
v19 <- load_variables(2019, "acs5", cache = TRUE)

##pull 2015-19 ACS place-level data from API using get_acs
census_2019 <- get_acs(
  geography = "place",
  year = 2019,
  survey = "acs5",
  output = "wide",
  geometry = FALSE,
  moe_sum = FALSE,
  key = census_api_key,
  variables = c(
    "B01003_001E", # pop_num,
    "B03002_006E", # pop_race_asian_num
    "B03002_004E", # pop_race_black_num
    "B03002_012E", # pop_race_hisp_num
    "B03002_003E", # pop_race_white_num
    "B25003H_002E", # homeowner_nonHisp_white_est
    "B25003H_001E", # count_nonHisp_white_est
    "B25003B_002E", # homeowner_Black_est
    "B25003B_001E", # count_Black_est
    "B25003D_002E", # homeowner_Asian_est
    "B25003D_001E", # count_Asian_est
    "B25003I_002E", # homeowner_Latino_est
    "B25003I_001E", # count_Latino_est
    "B25003_002E", #homeowner_est 
    "B25003_001E", #household_count_est
    "B19013B_001E", # median_household_income_Black
    "B19013H_001E", # median_household_income_white
    "B19013D_001E", # median_household_income_Asian
    "B19013I_001E", # median_household_income_Hispanic
    "B25077_001E" #median_value_own_occ

  )
) 
##create city size groupings
census_2019 <- census_2019 %>%
  mutate(size_type_update = case_when(
    B01003_001E >= 50000 & B01003_001E < 500000 ~ "Small and midsize cities  (50k-500k)",
    B01003_001E >= 500000 ~ "Large cities (500k +)"
  )) %>%
  filter(!is.na(size_type_update))

# Select variables needed for the analysis below
census_2019 <- census_2019 %>%
  select(
    geoid = "GEOID",
    NAME,
    pop_num =  "B01003_001E", 
    pop_race_asian_num = "B03002_006E", 
    pop_race_black_num = "B03002_004E", 
    pop_race_hisp_num = "B03002_012E", 
    pop_race_white_num  = "B03002_003E", 
    homeowner_nonHisp_white_est = "B25003H_002E", 
    count_nonHisp_white_est = "B25003H_001E", 
    homeowner_Black_est = "B25003B_002E", 
    count_Black_est = "B25003B_001E", 
    homeowner_Asian_est = "B25003D_002E", 
    count_Asian_est = "B25003D_001E", 
    homeowner_Latino_est = "B25003I_002E", 
    count_Latino_est = "B25003I_001E", 
    homeowner_est = "B25003_002E", #homeowner_est 
    household_count_est = "B25003_001E", #household_count_est
    median_household_income_Black = "B19013B_001E",
    median_household_income_white = "B19013H_001E",
    median_household_income_Asian =  "B19013D_001E", 
    median_household_income_Hispanic = "B19013I_001E", 
    median_value_own_occ = "B25077_001E", #
    "size_type_update"
  )
# Save as a data frame
census_2019 <- data.frame(census_2019)
```



```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=TRUE}
#### 2000 Decenial Census data
# Searching census codes
v2000 <- load_variables(2000, "sf1", cache = TRUE)
#write_csv(v2000, "v2000.csv")

#Creating a dataframe that has all the 2000 variables 
variables_df <- c(
     "P001001", #pop_2000
     "P008003", #pop_white_nh_2000
     "P008004", #pop_black_nh_2000
     "P008006", #pop_asian_nh_2000
     "P008010", #pop_hispanic_2000
    "H015I002",#	homeowner_nonHisp_white_est2000
    "H015I001",#	count_nonHisp_white_est2000
    "H015B002",#	homeowner_Black_est2000
    "H015B001",#	count_Black_est2000
    "H015D002",#	homeowner_Asian_est2000
    "H015D001", #	count_Asian_est2000
    "H015H002",#	homeowner_Latino_est2000
    "H015H001",#	count_Latino_est2000
    "H004002",#	homeowner_est2000
    "H004001" #	count_households_est2000
  )
# Pulling the 2000 variables from ACS
variables_2000_list <- map(variables_df, function(x) 
  {(variables_2000 <- get_decennial(
  geography = "place",
  year = 2000,
  sumfile = "sf1",
  output = "wide",
  key = census_api_key,
  variables = x))})

census_2000 <- reduce(variables_2000_list,left_join, by = c("GEOID", "NAME"))

# Select variables needed for the analysis below
census_2000 <- census_2000 %>%
  select(
    geoid = "GEOID",
    pop_2000 = "P001001", #
    pop_white_nh_2000 = "P008003", #
    pop_black_nh_2000 = "P008004", #
    pop_asian_nh_2000 = "P008006", #
    pop_hispanic_2000 = "P008010", #
   homeowner_nonHisp_white_est2000 = "H015I002",
    count_nonHisp_white_est2000= "H015I001",
    homeowner_Black_est2000 = "H015B002",#	
    count_Black_est2000 = "H015B001",#	
    homeowner_Asian_est2000 = "H015D002",#	
    count_Asian_est2000 = "H015D001", #	
    homeowner_Latino_est2000 = "H015H002",#	
    count_Latino_est2000 = "H015H001",#	
    homeowner_est2000 = "H004002",#	
    household_count_est2000 = "H004001" #	
  )
census_2000 <- census_2000 %>%
  mutate(homeowner_rate_nonHisp_white2000 = (homeowner_nonHisp_white_est2000 / count_nonHisp_white_est2000),
         homeowner_rate_Black2000 = (homeowner_Black_est2000 / count_Black_est2000),
         homeowner_rate_Asian2000 = (homeowner_Asian_est2000 / count_Asian_est2000),
         homeowner_rate_Latino2000 = (homeowner_Latino_est2000 / count_Latino_est2000),
         homeowner_rate2000 = (homeowner_est2000 / household_count_est2000)
         )


##Pull 2000 decennial Census median rent figures
median_hh_income_Black_2000 <-
  fromJSON(paste0("https://api.census.gov/data/2000/dec/sf3?get=P152B001,NAME&for=place:*&key=", census_api_key))%>%
  as_tibble() %>%
  transmute(
    median_hh_income_Black_2000 = V1,
    geoid = str_glue("{V3}{V4}")
  ) %>%
  filter(median_hh_income_Black_2000 != "P152B001")%>%
  mutate(
    median_hh_income_Black_2000 = as.numeric(median_hh_income_Black_2000)
  )
##Pull 2000 decennial Census median rent figures
median_hh_income_white_nh_2000 <-
  fromJSON(paste0("https://api.census.gov/data/2000/dec/sf3?get=P152I001,NAME&for=place:*&key=", census_api_key)) %>%
  as_tibble() %>%
  transmute(
    median_hh_income_white_nh_2000 = V1,
    geoid = str_glue("{V3}{V4}")
  ) %>%
  filter(median_hh_income_white_nh_2000 != "P152I001")%>%
  mutate(
    median_hh_income_white_nh_2000 = as.numeric(median_hh_income_white_nh_2000)
  )
##Pull 2000 decennial Census median rent figures
median_hh_income_Asian_2000 <-
  fromJSON(paste0("https://api.census.gov/data/2000/dec/sf3?get=P152D001,NAME&for=place:*&key=", census_api_key)) %>%
  as_tibble() %>%
  transmute(
    median_hh_income_Asian_2000 = V1,
    geoid = str_glue("{V3}{V4}")
  ) %>%
  filter(median_hh_income_Asian_2000 != "P152D001")%>%
  mutate(
    median_hh_income_Asian_2000 = as.numeric(median_hh_income_Asian_2000)
  )
##Pull 2000 decennial Census median rent figures
median_hh_income_hispanic_2000 <-
  fromJSON(paste0("https://api.census.gov/data/2000/dec/sf3?get=P152H001,NAME&for=place:*&key=", census_api_key)) %>%
  as_tibble() %>%
  transmute(
    median_hh_income_hispanic_2000 = V1,
    geoid = str_glue("{V3}{V4}")
  ) %>%
  filter(median_hh_income_hispanic_2000 != "P152H001")%>%
  mutate(
    median_hh_income_hispanic_2000 = as.numeric(median_hh_income_hispanic_2000)
  )
median_value_own_occ_2000 <-
  fromJSON(paste0("https://api.census.gov/data/2000/dec/sf3?get=H085001,NAME&for=place:*&key=", census_api_key)) %>%
  as_tibble() %>%
  transmute(
    median_value_own_occ_2000 = V1,
    geoid = str_glue("{V3}{V4}")
  ) %>%
  filter(median_value_own_occ_2000 != "H085001")%>%
  mutate(
    median_value_own_occ_2000 = as.numeric(median_value_own_occ_2000)
  )


# Join 2000 data
census_2000 <- census_2000 %>%
  full_join(median_hh_income_Black_2000, by = "geoid")%>%
  full_join(median_hh_income_white_nh_2000, by = "geoid")%>%
  full_join(median_hh_income_Asian_2000, by = "geoid")%>%
  full_join(median_hh_income_hispanic_2000, by = "geoid")%>%
  full_join(median_value_own_occ_2000, by = "geoid")

# Adjust for inflation using the CPi (The CPI is defined in "Define External Variables")
census_2000 <- census_2000 %>%
  mutate(median_hh_income_Black_2000_norm = CPI * median_hh_income_Black_2000,
         median_hh_income_white_nh_2000_norm = CPI * median_hh_income_white_nh_2000,
         median_hh_income_Asian_2000_norm = CPI * median_hh_income_Asian_2000,
         median_hh_income_hispanic_2000_norm = CPI * median_hh_income_hispanic_2000,
         median_value_own_occ_2000_norm = CPI * median_value_own_occ_2000)

```

```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=TRUE}
#Joining 2019 to 2000 data
homeownership_data <- census_2019 %>%
  full_join(census_2000, by = "geoid")%>%
  filter(size_type_update%in% c("Small and midsize cities  (50k-500k)", "Large cities (500k +)"))
``` 


```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=FALSE}
# Dropping cities that were not included in the 2000 census, cities outside the 50-states, and those that merged with their county jurisdiction to form a unified government over the period.
homeownership_data <- filter(homeownership_data, ! geoid %in% c("0464210", "0621230", "0637692", "0646842", "0812815", "1200410", "1245060", "1310944", "1319000", "1342425", "1349008", "1372122", "1373784", "1571550", "2148006", "2578972", "2483775", "5103320", "5367167", "1271625", "7276770", "7206593", "7214290", "7263820", "7210334", "7232522", "7252431" ))
```

     
```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=FALSE}

# Define placed as "principal city"  as "Principal" and those excluded as "suburbs". More on the Census designations here: https://www.census.gov/programs-surveys/metro-micro/about.html#:~:text=The%20largest%20city%20in%20each,concerning%20population%20size%20and%20employment.

principal_cities <- read_csv("metro_micro_principal_cities_2018_census.csv", n_max = 1268) %>%
  mutate(principal_city = TRUE)%>%
  mutate(
    format_zeros= str_pad(geoid, width = 7, side = "left", pad = "0") # Address leading zeros issue
  )%>%
  select(format_zeros, principal_city) %>%
  rename(geoid = format_zeros) 

# Join data, and create principal and suburbs flag
homeownership_data <- homeownership_data %>%
    left_join(principal_cities, by = "geoid")%>%
  mutate(place_type = case_when(
    principal_city == TRUE ~ "Principal",
    is.na(principal_city) ~ "Suburbs"
  )) %>%
  mutate(place_type_flag = case_when(
    principal_city == TRUE ~ 1,
    is.na(principal_city) ~ 0
  ))

## Create state IDs based on geoid
homeownership_data <- homeownership_data %>%
 mutate(stateid = substr(geoid, 1,2)
 )
```


```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=TRUE}
#### Summary Statsitics by city size
cities_size_homeownership<-
  homeownership_data %>%
  group_by(size_type_update) %>%
  summarise(
    pop_num = sum(pop_num), n = n(), homeowner_nonHisp_white_est = sum(homeowner_nonHisp_white_est), count_nonHisp_white_est = sum(count_nonHisp_white_est), homeowner_Black_est = sum(homeowner_Black_est), count_Black_est = sum(count_Black_est),homeowner_Asian_est = sum(homeowner_Asian_est), count_Asian_est = sum(count_Asian_est),homeowner_Latino_est = sum(homeowner_Latino_est), count_Latino_est = sum(count_Latino_est), homeowner_nonHisp_white_est2000 = sum(homeowner_nonHisp_white_est2000, na.rm = TRUE), count_nonHisp_white_est2000 = sum(count_nonHisp_white_est2000, na.rm = TRUE), homeowner_Black_est2000 = sum(homeowner_Black_est2000, na.rm = TRUE), count_Black_est2000 = sum(count_Black_est2000, na.rm = TRUE),homeowner_Asian_est2000 = sum(homeowner_Asian_est2000, na.rm = TRUE), count_Asian_est2000 = sum(count_Asian_est2000, na.rm = TRUE),homeowner_Latino_est2000 = sum(homeowner_Latino_est2000, na.rm = TRUE), count_Latino_est2000 = sum(count_Latino_est2000, na.rm = TRUE), pop_2000 = sum(pop_2000, na.rm = TRUE), pop_white_nh_2000 = sum(pop_white_nh_2000, na.rm = TRUE), pop_black_nh_2000 = sum(pop_black_nh_2000, na.rm = TRUE), pop_asian_nh_2000 = sum(pop_asian_nh_2000, na.rm = TRUE), pop_hispanic_2000 = sum(pop_hispanic_2000, na.rm = TRUE)
    ) %>%
  ungroup()%>%
  mutate(homeowner_rate_nonHisp_white = homeowner_nonHisp_white_est / count_nonHisp_white_est,
          homeowner_rate_Black = homeowner_Black_est / count_Black_est,
          homeowner_rate_Asian = homeowner_Asian_est / count_Asian_est,
          homeowner_rate_Latino = homeowner_Latino_est / count_Latino_est,
          homeowner_rate_nonHisp_white2000 = homeowner_nonHisp_white_est2000 / count_nonHisp_white_est2000,
          homeowner_rate_Black2000 =  homeowner_Black_est2000 / count_Black_est2000,
          homeowner_rate_Asian2000 = homeowner_Asian_est2000 / count_Asian_est2000,
          homeowner_rate_Latino2000 = homeowner_Latino_est2000 / count_Latino_est2000,
          ownership_change_nonHisp_white = homeowner_rate_nonHisp_white - homeowner_rate_nonHisp_white2000,
          ownership_change_Black = homeowner_rate_Black - homeowner_rate_Black2000,
          ownership_change_Asian = homeowner_rate_Asian - homeowner_rate_Asian2000,
          ownership_change_Latino = homeowner_rate_Latino - homeowner_rate_Latino2000,
          Black_white_gap = homeowner_rate_nonHisp_white - homeowner_rate_Black,
          Asian_white_gap = homeowner_rate_nonHisp_white - homeowner_rate_Asian,
          Latino_white_gap = homeowner_rate_nonHisp_white - homeowner_rate_Latino,
         )

```



```{r, echo=FALSE,message = FALSE, warning=FALSE, include=TRUE}
#### Format data for figures
homeownership_change_gg<-  cities_size_homeownership%>%
  transmute(size_type_update, ownership_change_nonHisp_white, ownership_change_Black, ownership_change_Asian, ownership_change_Latino)

chart_homeownership_change <-  melt(homeownership_change_gg, id.vars = c("size_type_update"), measure.vars = c("ownership_change_nonHisp_white", "ownership_change_Black", "ownership_change_Asian", "ownership_change_Latino"))

chart_homeownership_change <- chart_homeownership_change %>%
  mutate(variable = case_when(
             variable == "ownership_change_nonHisp_white" ~ "Non-Hispanic White",
             variable == "ownership_change_Black" ~ "Black",
             variable == "ownership_change_Asian" ~ "Asian",
             variable == "ownership_change_Latino" ~ "Hispanic"
           ))%>%mutate(
             value = value*100
           )

```



```{r, echo=FALSE,message = FALSE, warning=FALSE,  include=TRUE}
#### Figure 1: Change in Homeownership Rates by Race and Ethnicity, 2000 - 2019 
figure_1 <- ggplot(chart_homeownership_change, 
                   aes(size_type_update, value,fill=interaction(variable), label = scales::number(value, .1), tooltip =scales::percent(value, .1), 
                       text = sprintf("City size category: %s<br>Race and ethnicty: %s<br>Percentage point change in homeownership rate, 2000- 2019: %s", 
      size_type_update, variable, 
      scales::comma(value, accuracy = .1) 
      )
                         )) +
  geom_bar(stat='identity', width = .75, position='dodge') + #add custom data label notation
   annotate(geom = "text", x = .7, y = 5.5, label = "5.2", hjust = "left", size = 3.4)+ #add custom data label notation
   annotate(geom = "text", x = .9, y = -4.8, label = "-4.4", hjust = "left", size = 3.4)+ #add custom data label notation
   annotate(geom = "text", x = 1.1, y = 2.75, label = "2.4", hjust = "left", size = 3.4)+ #add custom data label notation
   annotate(geom = "text", x = 1.27, y = -.9, label = "-0.6", hjust = "left", size = 3.4)+ #add custom data label notation
   annotate(geom = "text", x = 1.72, y = 6.08, label = "5.8", hjust = "left", size = 3.4)+#add custom data label notation
  annotate(geom = "text", x = 1.89, y = -4.8, label = "-4.4", hjust = "left", size = 3.4)+#add custom data label notation
    annotate(geom = "text", x = 2.1, y = .4, label = "0.2", hjust = "left", size = 3.4)+#add custom data label notation
      annotate(geom = "text", x = 2.28, y = -1.28, label = "-0.9", hjust = "left", size = 3.4)+#add custom data label notation

  theme_bw() + theme(axis.text.x = element_text( hjust=1)) +
  scale_fill_manual(values =  c("#1f687e","#F1D8D4","#a6e0e0", "#904f35"))+ #add color fill 
    scale_y_continuous(labels = scales::number, breaks = pretty_breaks())+  #format labels
  labs(
    x = "",
    y = "Percentage point change in homeownership",
    color = NULL
  )+
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 8), #format title
    axis.title.y = element_text(size = 8), #format title
    axis.text.x = element_text(hjust = 1, size = 8), # tilts the x-axis text
    axis.text.y = element_text(size = 8), # tilts the x-axis text
    legend.title = element_blank(), #format legend
    legend.text = element_text(size = 10), #format legend
    legend.position = "top", #format legend
    plot.caption = element_text(hjust = 0, face = "italic", size = 10), # format the source caption
    plot.title.position = "plot",
    plot.caption.position = "plot"
  )
##make the figure interactive 
plot_interactive_figure_1 <- ggplotly(figure_1, tooltip = 'text')%>%
 layout(annotations = 
 list(x = 1, y = -0.2, text = "<i>Source: 2015-2019 5-year American Community Survey and New York State ERAP Application Figures as of 8/31/21 (NY OTDA), NYU Furman Center.</i>", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(family = "verdana", size=8)))

plot_interactive_figure_1

#Export as HTML
#saveWidget(plot_interactive_figure_1, "/Users/carlhedman/Documents/NYU/Race and homeownership blog/Final figures/Webfiles/plot_interactive_figure_1.html")
```


```{r, echo=FALSE,fig.keep='all', results='hide', warning=FALSE, include=FALSE}
#### Pull in shapefiles for maps
##Loop to create all states
for (i in state.abb) {
nam <- paste("pl_", i, sep = "")
assign(nam,
       places(state=i, cb = TRUE, year=2018)%>%
  mutate(geoid = GEOID))
}
##pull in the DC shapefile
pl_DC <- places(state=11, cb = TRUE, year=2018)%>% 
  mutate(geoid = GEOID)

##Bind all state place level data to create national map shapefile
pl_all <- rbind (pl_AK, pl_AL, pl_AR, pl_AZ, pl_CA, pl_CO, pl_CT, pl_DE, pl_DC, pl_FL, pl_GA, pl_HI, pl_IA, pl_ID, pl_IL, pl_IN, pl_KS, pl_KY, pl_LA, pl_MA, pl_MD, pl_ME, pl_MI, pl_MN, pl_MO, pl_MS, pl_MT, pl_NC, pl_ND, pl_NE, pl_NH, pl_NJ, pl_NM, pl_NV, pl_NY, pl_OH, pl_OK, pl_OR, pl_PA, pl_RI, pl_SC, pl_SD, pl_TN, pl_TX, pl_UT, pl_VA, pl_VT, pl_WA, pl_WI, pl_WY )

##Join to data set
place_data <- pl_all %>%
  inner_join(homeownership_data, by = "geoid")

##Create color scheme
pal <- colorFactor(
  palette = c('#1f687e', '#d83d22', '#904f35', '#c8e9f1'),
  levels= place_data$Black_ownership_change
)
 
```


```{r, echo=TRUE,fig.keep='all', results='hide', warning=FALSE, include=FALSE}
#### Create city map choropleth bins and map labels
##note: table code broadly uses HTML notation to construct 
labels_table <- sprintf("<strong>%s</strong><br/>
<i>%s</i><br/>
Population in 2019: <strong>%s</strong><br/>
Median value of owner-occupied homes, 2019: <strong>$%s</strong><br/>
    <style>
      table {
        border-collapse: collapse;
      }
      td,
      th {
        padding: 5px;
        border-bottom: 1.5px solid #1c87c9;
      }
    </style>
<table>
  <tr>
    <td>&nbsp;</td> #specify row  labels
    <th>Black</td> #specify row  labels
    <th>White</td> #specify row  labels
    <th>Asian</td> #specify row  labels
    <th>Hispanic</td> #specify row  labels
  </tr>
    <tr>
    <td>Share of Population, `19</td> #specify column labels
    <td>%1.1f%%</td> #create cell for data 
    <td>%1.1f%%</td>  #create cell for data 
    <td>%1.1f%%</td> #create cell for data 
    <td>%1.1f%%</td> #create cell for data  
  </tr>
    <tr>
    <td>Homeownership Rate, `19</td> #specify column labels
    <td>%1.1f%%</td>#create cell for data 
    <td>%1.1f%%</td> #create cell for data 
    <td>%1.1f%%</td> #create cell for data 
    <td>%1.1f%%</td>#create cell for data 
  </tr>
  <tr>
    <td>Change in Homeownership Rate<br>Percentage Points, '00-'19</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
  </tr>
  <tr>
    <td>Change in Number of Homeowners<br>Numeric, '00-'19</td>
    <td>%+s</td>
    <td>%+s</td>
    <td>%+s</td>
    <td>%+s</td>
  </tr>
    <tr>
    <td>Median Household Income, `19</td>
    <td>$%s</td>
    <td>$%s</td>
    <td>$%s</td>
    <td>$%s</td>
  </tr>
</table>
                        ",
  place_data$NAME.y, #include the data in the order specified above
  place_data$Black_ownership_change_long,  #include the data in the order specified above
  prettyNum(place_data$pop_num, big.mark = ","), #include the data in the order specified above, include formatting
  prettyNum(place_data$median_value_own_occ, big.mark = ","), #include the data in the order specified above, include formatting

  place_data$share_pop_black*100, #include the data in the order specified above, include formatting
  place_data$share_pop_white*100,
  place_data$share_pop_asian*100,
  place_data$share_pop_hisp*100,

  place_data$homeowner_rate_Black*100, #include the data in the order specified above, include formatting
  place_data$homeowner_rate_nonHisp_white*100, 
  place_data$homeowner_rate_Asian*100,  
  place_data$homeowner_rate_Latino*100,  

  place_data$ownership_per_change_Black*100, #include the data in the order specified above, include formatting
  place_data$ownership_per_change_nonHisp_white*100,  
  place_data$ownership_per_change_Asian*100, 
  place_data$ownership_per_change_Latino*100,

  prettyNum(place_data$ownership_count_change_Black, big.mark = ","), #include the data in the order specified above, include formatting
  prettyNum(place_data$ownership_count_change_nonHisp_white, big.mark = ","),
  prettyNum(place_data$ownership_count_change_Asian, big.mark = ","),
  prettyNum(place_data$ownership_count_change_Latino, big.mark = ","),

  prettyNum(place_data$median_household_income_Black, big.mark = ","),#include the data in the order specified above, include formatting
  prettyNum(place_data$median_household_income_white, big.mark = ","),
  prettyNum(place_data$median_household_income_Asian, big.mark = ","),
  prettyNum(place_data$median_household_income_Hispanic, big.mark = ",")
) %>% lapply(htmltools::HTML)

##same structure as above
labels_popup <- sprintf("<strong>%s</strong><br/>
<i>%s</i><br/>
Population in 2019: <strong>%s</strong><br/>
Median value of owner-occupied homes, 2019: <strong>$%s</strong><br/>
    <style>
      table {
        border-collapse: collapse;
      }
      td,
      th {
        padding: 5px;
        border-bottom: 1.5px solid #1c87c9;
      }
    </style>
<table>
  <tr>
    <td>&nbsp;</td>
    <th>Black</td>
    <th>White</td>
    <th>Asian</td>
    <th>Hispanic</td>
  </tr>
    <tr>
    <td>Share of Population, `19</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
  </tr>
    <tr>
    <td>Homeownership Rate, `19</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
    <td>%1.1f%%</td>
  </tr>
  <tr>
    <td>Change in Homeownership Rate<br>Percentage Points, '00-'19</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
    <td>%+1.1f</td>
  </tr>
  <tr>
    <td>Change in Number of Homeowners<br>Numeric, '00-'19</td>
    <td>%+s</td>
    <td>%+s</td>
    <td>%+s</td>
    <td>%+s</td>
  </tr>
    <tr>
    <td>Median Household Income, `19</td>
    <td>$%s</td>
    <td>$%s</td>
    <td>$%s</td>
    <td>$%s</td>
  </tr>
</table>
                        ",
  place_data$NAME.y, 
  place_data$Black_ownership_change_long,
  prettyNum(place_data$pop_num, big.mark = ","),
  prettyNum(place_data$median_value_own_occ, big.mark = ","),

  place_data$share_pop_black*100,
  place_data$share_pop_white*100,
  place_data$share_pop_asian*100,
  place_data$share_pop_hisp*100,

  place_data$homeowner_rate_Black*100, 
  place_data$homeowner_rate_nonHisp_white*100, 
  place_data$homeowner_rate_Asian*100,  
  place_data$homeowner_rate_Latino*100,  

  place_data$ownership_per_change_Black*100, 
  place_data$ownership_per_change_nonHisp_white*100,  
  place_data$ownership_per_change_Asian*100, 
  place_data$ownership_per_change_Latino*100,

  prettyNum(place_data$ownership_count_change_Black, big.mark = ","),
  prettyNum(place_data$ownership_count_change_nonHisp_white, big.mark = ","),
  prettyNum(place_data$ownership_count_change_Asian, big.mark = ","),
  prettyNum(place_data$ownership_count_change_Latino, big.mark = ","),

  prettyNum(place_data$median_household_income_Black, big.mark = ","),
  prettyNum(place_data$median_household_income_white, big.mark = ","),
  prettyNum(place_data$median_household_income_Asian, big.mark = ","),
  prettyNum(place_data$median_household_income_Hispanic, big.mark = ",")
) %>% lapply(htmltools::HTML)
```


#### Figure 2: City-Level Change in Black Homeownership: Michigan
*Zoom in and hover over the Zip Code Tabulation Areas to explore variations across the state*
```{r, echo=FALSE,message = FALSE, warning=FALSE,  include=TRUE}
MI_map <- leaflet(data = place_data) %>% addProviderTiles(providers$CartoDB.Positron) %>% #specify data source and shapefiles
    setView(lng = -84.5, lat = 42.3, zoom = 8.4)%>% #set Zoom and view lat/long
  addPolygons(fillColor = ~pal(Black_ownership_change), #add fill color scheme
  weight = 1, #add line weight
  opacity = 1, #add line opacity
color = "#444444", #add line color
  dashArray = "0", #add line style
  fillOpacity = 0.8, #add fill opacity
  highlight = highlightOptions( #mouseover/highlight parameters
    weight = 2,
    color = "white",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
   label = labels_table, #add mouseover labels
  popup = labels_popup, #add click popup labels
  popupOptions = popupOptions(maxWidth = 900, closeOnClick = TRUE), #add click popup options
   labelOptions = labelOptions(  #add label  options
     style = list("font-weight" = "normal", padding = "3px 8px"),
     textsize = "10px",
     direction = "auto")
)%>% 
  addLegend(pal = pal, values = ~Black_ownership_change, opacity = 0.7, title = "Numeric and Perc. Point Change<br>in Black Homeownership, 00`-19`",
  position = "bottomright")  #add legend, specify features

MI_map
#export interactive map as HMTL 
#mapshot(MI_map, url =  "/Users/carlhedman/Documents/NYU/Race and homeownership blog/Final figures/Webfiles/MI_map.html")
```
*Source: 2015-2019 5-year American Community Survey and 2000 Decennial sf3 Census, Housing Solutions Lab.*

<br>
<br>

#### Figure 3: City-Level Change in Black Homeownership: Texas
*Zoom in and hover over the Zip Code Tabulation Areas to explore variations across the state*
```{r, echo=FALSE,message = FALSE, warning=FALSE, include=TRUE}
TX_map <- leaflet(data = place_data) %>% addProviderTiles(providers$CartoDB.Positron) %>% #specify data source and shapefiles
    setView(lng = -96.90, lat = 32.54, zoom = 7)%>%
  addPolygons(fillColor = ~pal(Black_ownership_change),
  weight = 1,
  opacity = 1,
color = "#444444",
  dashArray = "0",
  fillOpacity = 0.8,
  highlight = highlightOptions(
    weight = 2,
    color = "white",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
   label = labels_table,
  popup = labels_popup,
  popupOptions = popupOptions(maxWidth = 1000, closeOnClick = TRUE),
   labelOptions = labelOptions(
     style = list("font-weight" = "normal", padding = "3px 8px"),
     textsize = "10px",
     direction = "auto")
)%>% 
  addLegend(pal = pal, values = ~Black_ownership_change, opacity = 0.7, title = "Numeric and Perc. Point Change<br>in Black Homeownership, 00`-19`",
  position = "bottomright")

TX_map
#export interactive map as HMTL 
#mapshot(TX_map, url =  "/Users/carlhedman/Documents/NYU/Race and homeownership blog/Final figures/Webfiles/TX_map.html")
```
*Source: 2015-2019 5-year American Community Survey and 2000 Decennial sf3 Census, Housing Solutions Lab.*


#### Figure 4: City-Level Change in Black Homeownership: National
*Zoom in and hover over the Zip Code Tabulation Areas to explore variations across the state*
```{r, echo=FALSE,message = FALSE, warning=FALSE, include=TRUE}
USA_map <- leaflet(data = place_data) %>% addProviderTiles(providers$CartoDB.Positron) %>%
    setView(lng = -96.90, lat = 32.54, zoom = 4)%>%
  addPolygons(fillColor = ~pal(Black_ownership_change),
  group = "name",
  weight = 1,
  opacity = 1,
color = "#444444",
  dashArray = "0",
  fillOpacity = 0.8,
  highlight = highlightOptions(
    weight = 2,
    color = "white",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
   label = labels_table,
  popup = labels_popup,
 popupOptions = popupOptions(maxWidth = 1000, closeOnClick = TRUE),
   labelOptions = labelOptions(
     style = list("font-weight" = "normal", padding = "3px 8px"),
     textsize = "10px",
     direction = "auto")
)%>%
  addLegend(pal = pal, values = ~Black_ownership_change, opacity = 0.7, title = "Numeric and Perc. Point Change<br>in Black Homeownership, 00`-19`",
  position = "bottomright")%>%
   addResetMapButton() %>%
addSearchOSM(options = searchOptions(autoCollapse = TRUE, minLength = 2, zoom = 8, hideMarkerOnCollapse = TRUE)) ##add search bar

USA_map

#export interactive map as HMTL 
#mapshot(USA_map, url =  "/Users/carlhedman/Documents/NYU/Race and homeownership blog/Final figures/Webfiles/USA_map.html")

```

